New lower bounds for Schur and weak Schur numbers

### 2021-10-13 article/weakschur.tex
modifs ws + les there is / are arrivent

In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers \cite{AbbottHanson}. By analogy with the previous section,
we then introduce \textit{WS-templates}, standing for \textit{"weak Schur templates"}, as well as an associated sequence \(\WS^+(n)\). 
We find templates and use them to provide new lower bounds for weak Schur numbers. Finally, we give a short explanation for 
the new lower bound \(\WS(6) \geqslant 646\) which was not directly obtained with a template contrary to the other lower bounds 
given in this article.

@@

\begin{theorem}
\label{thm:WS-ab}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \({[\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]}\) into \(n+k\)
weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

In particular, by setting \(q = \WS (n)\) and \(p = S(k)\) in Theorem \ref{thm:WS-ab}, one obtains the following corollary.

\begin{corollary}
\label{cor:ineqWS-S}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } \WS (n+k) \geqslant S(k) \left (\WS (n) + \left \lceil \displaystyle \frac{\WS (n)}{2}
\right \rceil +1 \right) + \WS (n)\)
\end{corollary}

This can be seen as an equivalent for weak Schur numbers of Abott and Hanson's construction for Schur numbers. This formula includes
the results of Rowley \cite{RowleyWS} as a special case. For \(n>2\), this formula does not give new lower bounds.

\begin{remark}
The inequality from Corollary \ref{cor:ineqWS-S} can be improved by adding 1 to the lower bound if \(\WS (n)\) is odd (more generally if \(q\) is
odd in Theorem \ref{thm:WS-ab}). However, it lengthens the proof and it is never useful in practice.
\end{remark}

Given that Theorem \ref{thm:WS-ab} will appear as a particular case of a more general theorem after the introduction of
templates for weak Schur numbers, we only give here an intuitive explanation of the demonstration; a formal
\hyperref[PreuveThm]{proof} using templates for weak Schur numbers is provided in Subsection \ref{ConstructionWS}.

@@

\begin{itemize}
	\item \(\mathcal{T}\) (the "tail"): the integers from 1 to \(q\). NB: this is line number 0.
	\item \(\mathcal{R}\) (the "rows"): the integers in columns \(-l\) to \(+r\) (excluding  \(\mathcal{T}\)).
	\item \(\mathcal{C}\) (the "columns"): the integers in the last \(w\) columns (excluding  \(\mathcal{T}\)).
\end{itemize}

Like S-templates, \(\mathcal{R}\) and \(\mathcal{C}\) play the role of security zones for each other. Note that with
this numbering of columns, the column of the sum of two numbers is the only integer in \([\![-l,q]\!]\) equal to two the
sum of the columns modulo \(a\).

@@

\noindent \underline{\textbf{\(\mathcal{R}\) block}}
\par
In this block,  we use the colors \(n + 1, ..., n + k\). We color an integer \(x\) according to its line number (written \(\lambda(x)\)).
For every \(x \in \mathcal{R}\), we color \(x\) with \(n + c\) where \(c\) is the color of \(\lambda(x)\) in the sum-free coloring of  \([\![1,p]\!]\).
Let \((x, y) \in \mathcal{R}^2\). The cases are twofold.

\begin{itemize}
	\item \underline{\(\lambda(x+y) = \lambda(x) + \lambda(y)\)} \\
	In this case, we use the sum-free property of the coloring of \([\![1,p]\!]\) (in block \(\mathcal{C}\), we only
	use colors \(1, ..., n\)).
	\item \underline{\(\lambda(x+y) \neq \lambda(x) + \lambda(y)\)} \\
	In this case, we do not have information about the color of \(\lambda(x+y)\). Thereby, we want to have
	\(x+y \in \mathcal{C}\). A simple solution is to limit the horizontal movement made by summing the two integers.
	That is, when the sum's line number is different from \(\lambda(x) + \lambda(y)\), not to move too far from \((\lambda(x)
	+ \lambda(y)) a\) so that \(x+y\) stays in \(\mathcal{C}\). There, the maximal displacement to the left (resp.
	to the right) is \(2l\) (resp. \(2r\)). Not crossing entirely \(\mathcal{C}\) by going to the left is then expressed
	as \(-2l > -a + r\). Likewise, not going to far to the right is expressed as \(2r < a - l\). It can then be written
	as \(\max(l, r) \leqslant w\).
\end{itemize}

\noindent \underline{\textbf{\(\mathcal{C}\) block}}
\par
In this block,  we use colors \(1, ..., n\). We color an integer \(x\) according to its column number, denoted by \(\tilde{\pi}(x)\). It is linked to the
projection on the first line, denoted by \(\pi\), by the relation \(\tilde{\pi}(x) = \pi(x) - a\). A simple solution is to color \(x\) with the same color
as \(\tilde{\pi}(x)\) in the weakly sum-free coloring of \([\![1,q]\!]\). As long as \(2q \leqslant a + r\) (not going two far to the right) and there
is no \(x \in \tilde{\pi}(\mathcal{C})\) such that \(2x \in \tilde{\pi}(\mathcal{C})\) (so that we do not have a sum in \(\mathcal{C}\) when
taking two numbers in the same column), the colors \(1, ..., n\) are sum-free.

In particular, taking \(w = l = \left\lceil \displaystyle \frac{q}{2} \right\rceil\) and \(r = \left\lfloor \displaystyle \frac{q}{2} \right\rfloor\) works, thus obtaining the above theorem.\\
\par
As in the previous section, we now introduce WS-templates and the sequence \(\WS^+\) in order to generalize the above construction.

@@

Note that the special color \(n\) is not necessarily the last color by order of appearance, any color can play this role.
We now introduce the number \(\WS^+(n)\) that plays the same role as \(S^+(n)\) for S-templates.
However, WS-templates being more sophisticated than S-templates, the definition of \(\WS^+(n)\)
is slighty more complicated.

@@

Having established these properties, we now have everything we need to state and prove our main result. We proceed to do so
in the next subsection.

@@

Theorem \ref{thm:Stemp} provides a way to extend a S-template into a larger Schur partitions using other Schur partitions. Likewise, a WS-template can 
be extended into larger weak Schur partitions by using Schur partitions. This is the object of the following theorem.

\begin{theorem}
\label{thm:WStemp}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,p]\!]\)
into \(k\) sum-free subsets and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there exists a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

In particular, by setting \(p = S(k)\) and \(a = \WS^+(n + 1)\) in Theorem \ref{thm:WStemp}, the next corollary follows.

\begin{remark}
In the S-template construction for Schur numbers, the additive constant comes from the fact that the special color does
not necessarily appear right at the begining of the repeating pattern. Likewise, \(b_{max}\) can actually be replaced by
\[
\max_{b \in \mathbb{N}^*} \left\{\min (A_{n+1} \backslash [\![1, b ]\!]) - 1~|~ \WS_b^+(n+1) = \WS^+(n+1) \right\}.
\]
\end{remark}

\begin{proof}[\textsc{Proof.}]
\begin{sloppypar}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) and \((p,k) \in (\mathbb{N}^*)^2\). Denote by \(f\) the coloring
associated to the \(b\)-WS-template  and \(g\) the one associated to the sum-free partition of \([\![1,p]\!]\); where
\({f : [\![1, a + b]\!] \longrightarrow [\![1,n+1]\!]}\) and \({g : [\![1, p]\!]  \longrightarrow [\![1, k]\!]}\). Moreover, assume
that the sum-free coloring of \([\![1, p]\!]\) is ordered.
\end{sloppypar}

@@

Split \([\![1, p a + b]\!]\) into three subsets. \\
NB: To keep the notation short, the restriction to \([\![b + 1, p a + b]\!]\) of \(\pi\) defined in Subsection \ref{DefinitionWS+}
is denoted by \(\pi\) in the hereunder equations.

@@

A new coloring \(h\) is defined as follows:
\[
\begin{array}{c c c l}
	h : & [\![1, p a + b]\!] & \longrightarrow & [\![1,n+k]\!] \\
	& x & \longmapsto &
	\left\{ \begin{array}{l l}
		f(x) & \text{if}~x \in \mathcal{T} \\
		f(\pi(x)) & \text{if}~x \in \mathcal{C} \\
		n + g(\lambda(x)) & \text{if}~x \in \mathcal{R}
	\end{array} \right.
\end{array}
\]

Function \(h\) is well defined since \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) is a partition of  \([\![1, p a + b]\!]\).
We now prove that \(h\) is a weakly sum-free coloring. Let \(x,y \in [\![1, p a + b]\!]\) be such that \(x \neq y\),
\(h(x) = h(y)\) and \(x+y \leqslant p a+ b\). We claim that \(h(x+y) \neq h(x)\). Nine cases are to be distinguished
according to the subsets \((\mathcal{T}, \mathcal{C}, \mathcal{R})\) to which \(x\) and \(y\) belong. It is
sufficient to check only six cases out of nine since \(x\) and \(y\) play symmetric roles. \\

@@

\noindent \underline{\textbf{Case 2:} \((x,y) \in \mathcal{T} \times \mathcal{C}\)}
\par
Given that \(h(x) = h(y)\) and by definition of \(h\), \(f(x) = f(\pi(y))\). Besides, \(f(\pi(y)) \leqslant n\) since
\(y \in \mathcal{C}\). Two cases are to be distinguished according to the value of \(x + \pi(y)\).
\begin{itemize}
\item \begin{sloppypar}
	If \(x + \pi(y) \leqslant a + b\) then \(f(x + \pi(y)) = f(\pi(x + y))\) (Proposition \ref{prop3}). Given that \(f\) is
	a weakly sum-free coloring, \(f(x+\pi(y)) \neq f(x)\) since \(f(x)=f(\pi(y))\) and \(x \neq \pi(y)\) since
	\({x \leqslant b < \pi(y)}\).
	\end{sloppypar}
\item \begin{sloppypar}
	If \(x+\pi(y)> a+b\) then given that \(f\) is a WS-template and since \({f(x) = f(\pi(y)) \leqslant n}\),
	\({f(\pi(x+\pi(y))) \neq f(x)}\).~Furthermore \({f(\pi(x+\pi(y))) = f(\pi(x+y))}\) (Proposition \ref{prop2}), such that
	\({f(\pi(x+ y)) \neq f(x)}\).
	\end{sloppypar}
\end{itemize}
\par
Hence in both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

\noindent \underline{\textbf{Case 3:} \((x,y) \in \mathcal{T} \times \mathcal{R}\)}
\par
Necessarily \(h(x) = h(y) = n + 1\). Two cases are to be distinguished according to the value of \(\lambda(x+y)\).
\begin{itemize}
\item If \(\lambda(y)=\lambda(x+y)\) then \(\pi(x + y) = x + \pi(y)\) (Proposition \ref{prop6}). By definition of
	\(h\), \(f(x) = f(\pi(y))\). Given that \(f\) is a weakly sum-free coloring, \(f(x + \pi(y)) \neq f(x)\) since
	\(f(x) = f(\pi(y))\) and \(x \neq \pi(y)\) since \({x \leqslant b < \pi(y)}\). Hence \(h(x + y) \neq h(x)\).
\item If \(\lambda(y) \neq \lambda(x + y)\) then \(\lambda(x + y) = \lambda(y) + 1\) since \(x \leqslant b < a\).
	Besides, \(n + 1 = h(y) = n +  g(\lambda(y))\). Hence \(g(\lambda(y)) = 1\). Moreover \(g(1) = 1\) since \(g\)
	is an ordered coloring. Therefore, given that \(g\) is sum-free, \(g(\lambda(y) + 1) \neq 1\). If \(\pi(x + y) \in
	A_{n + 1}\) then \(h(x + y) = n + g(\lambda(x + y)) \neq n + 1\). Otherwise, \(h(x + y) \leqslant n\). Hence
	in both cases \(h(x + y) \neq h(x)\).
\end{itemize}

\noindent \underline{\textbf{Case 4:} \((x,y) \in \mathcal{C}^2\)}
\par
By definition of \(h\) and since \(h(x)=h(y)\), \(f(\pi(x)) = f(\pi(y))\). Two cases are to be distinguished according
to the value of \(\pi(x)+\pi(y)\).
\begin{itemize}
\item If \(\pi(x) + \pi(y) \leqslant a+b\) then \(\pi(x)+\pi(y) = \pi(x + y)\). Hence \(f(\pi(x + y)) \neq f(\pi(x))\) since
	\(f\) is sum-free for \(x>b\)
\item \begin{sloppypar}
	If \(\pi(x)+\pi(y)>a+b\) then given that \(f\) is a WS-template, \({f(\pi(\pi(x)+\pi(y))) \neq f(\pi(x))}\) since
	\({f(\pi(x)) = f(\pi(y))}\). Besides,  \({f(\pi(\pi(x)+\pi(y))) = f(\pi(x + y))}\) (Proposition \ref{prop4}). Hence \({f(\pi(x + y))
	\neq  f(\pi(x))}\).
	\end{sloppypar}
\end{itemize}
\par
Hence in both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

@@

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the WS property inherited
from both the S-template and the WS-template. \\
\end{proof}

Theorem \ref{thm:WS-ab} can be seen as a particular case of WS-template in the same way Abott and Hanson's 
construction \cite{AbbottHanson} can be seen as a particular case of S-template.

@@

@@

\begin{proposition}
Let \((b, k, a) \in (\mathbb{N}^*)^3\) and let \(f\) be a coloring associated to a \(b\)-WS-template with width \(p\) and \(k\) colors. Let
\(c \in \mathbb{N}\) and assume there there exists a coloring \(g\) of \([\![b + 1, b + c]\!]\) with \(k\) colors such that for all \(c \in [\![1, k]\!]\),

@@

The WS-templates can actually be fine-tuned further. However, it only gives minor improvements (most likely only an additive
constant) at the cost of dramatically increasing the size of the search space. Therefore, it does not seem relevant to
use this sophistications given that we could not even find good WS-templates with five colors using a computer (here good
means better than those obtain by combining smaller templates).

\par These modifications work as follows. One may notice that the first row (excluding the "tail") has constraints that other rows
do not have because of the tail, especially if the special color appears in the tail as well. Thus allowing to have a coloring on the
first row different from the coloring of the other rows would weaken the constraints. Actually, one may even go further by
noticing that on the one hand the first (ordered) color of the sum-free partition used for the extension procedure has more
more constraints than the other colors of the sum-free partition since the first row is of this color and is more constrained than
the other rows, but that on the other hand it has more degrees of freedom than the other colors of the sum-free partition since in
the sum-free partition there cannot be two consecutive numbers of this color. As a result, it removes  some constraints imposed
 by the first row on the other rows.

\par
To sum up, one can look for a generalised WS-template that uses a special coloring for the tail and the first row, a
coloring dedicated to the rows whose number is not 1 but is in the first color in the sum-free partition, a coloring for all
the other rows and a special coloring for the last numbers (as previously explained for the improvement of the additive
constant of WS-templates).

@@

We produced WS-templates using a SAT solver, hence providing lower bound on \(WS^+\) and inequalities
of the type \(WS(n+k) \geqslant a S(n) + b\). We sought templates providing the greatest value of \((a, b)\) (in the 
lexicographic order). Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
Here are the inequalities given by the current best WS-templates. The \hyperref[WS-templates]{template} 
corresponding to the third inequality can be found in the appendix.

@@

However, trying out all of the 2\,447\,113\,088 Schur number 5-partitions \cite{Heule2017} one by one would not result in a 
reasonable computation time: it is necessary to test the construction on several partitions at once. We encoded the problem as a 
satisfiability problem and used the 1616 backdoors that were used in \cite{Heule2017} in order to encode a group of partitions in 
a compact and efficient way. Among all of the 1616 backdoors, only the 911\textsuperscript{th} backdoor gave a weakly sum-free 
partition of length 643. This backdoor gave a weakly sum-free partition of length 646 and cannot give a weakly sum-free 
partition of length 647. A weakly sum-free partition of length 646 can be found in the \hyperref[WS(6)]{appendix}.

>>>>

We generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS}. We give an analogous for weak Schur numbers of Abbott 
and Hanson's construction for Schur numbers \cite{AbbottHanson}. Then, as in the previous section, we introduce \textit{WS-templates}, 
standing for \textit{"weak Schur templates"}, as well as an associated sequence \(\WS^+(n)\). We find templates and use them to provide 
new lower bounds for weak Schur numbers. Finally, we give a short explanation for the new lower bound \(\WS(6) \geqslant 646\) which 
was not directly obtained with a template, contrary to the other lower bounds given in this article.

@@

\begin{theorem}
\label{thm:WS-ab}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there are a weakly sum-free \(n\)-partition of \([\![1,q]\!]\) and
a sum-frree \(k\)-partition of \([\![1,p]\!]\) then there is a partition of \({[\![1,p(q+\left \lceil \frac{q}{2} \right 
\rceil + 1)+q]\!]}\) into \(n+k\) weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

Given that Theorem \ref{thm:WS-ab} will appear as a particular case of a more general theorem after the introduction of
templates for weak Schur numbers, we only outline the demonstration; a \hyperref[PreuveThm]{proof} using templates 
for weak Schur numbers is in Subsection \ref{ConstructionWS}.

@@

\begin{itemize}
	\item \(\mathcal{T}\) (the "tail"): In the row indexed by 0: numbers from \([\![1,q]\!]\).
	\item \(\mathcal{R}\) (the "rows"): the integers in columns \(-l\) to \(+r\).
	\item \(\mathcal{C}\) (the "columns"): the integers in the last \(w\) columns.
\end{itemize}

Note that with this numbering of columns, the column of the sum of two numbers is the only integer in \([\![-l,q]\!]\) equal to the
sum of the columns modulo \(a\).

@@

\noindent \underline{\textbf{\(\mathcal{R}\) block}}
\par
We use the colors \(n + 1, ..., n + k\) in this block. We color an integer \(x\) according to its line number \(\lambda(x)\).
For any \(x \in \mathcal{R}\), we color \(x\) with \(n + c\) where \(c\) is the color of \(\lambda(x)\) in the sum-free coloring of  \([\![1,p]\!]\).
Let \((x, y) \in \mathcal{R}^2\). The cases are twofold.

\begin{itemize}
	\item \underline{\(\lambda(x+y) = \lambda(x) + \lambda(y)\)} \\
	If \(x + y\in \mathcal{R}\), we use the sum-free property of the coloring of \([\![1,p]\!]\). Otherwise, \(x + y\in \mathcal{C}\) which 
	is fine because we only use colors \(1, ..., n\) in block \(\mathcal{C}\).
	\item \underline{\(\lambda(x+y) \neq \lambda(x) + \lambda(y)\)} \\
	In this case, we do not have information about the color of \(\lambda(x+y)\). Thereby, we want to have
	\(x+y \in \mathcal{C}\). A simple solution is to limit the horizontal displacement made by summing the two integers.
	That is, when the sum's row number is different from \(\lambda(x) + \lambda(y)\), to upper bound 
	\(|x + y - (\lambda(x) + \lambda(y)) a|\) so that \(x+y\) stays in \(\mathcal{C}\). There, the maximal displacement 
	to the left (resp. to the right) is \(2l\) (resp. \(2r\)). Not crossing entirely \(\mathcal{C}\) by going to the left is then 
	expressed as \(-2l > -a + r\). Likewise, not going to far to the right is expressed as \(2r < a - l\). It can then be written
	as \(\max(l, r) \leqslant w\).
\end{itemize}

\noindent \underline{\textbf{\(\mathcal{C}\) block}}
\par
In this block, we use colors \(1, ..., n\). We color an integer \(x\) according to its projection on the first row \(\pi(x)\) offeset by \(a\), 
that is its column number. A simple solution is to color \(x\) with the same color as \(\pi(x) - a\) in the weakly sum-free coloring of 
\([\![1,q]\!]\). As long as \(2q \leqslant a + r\) (not going two far to the right) and there is no \(x \in \pi(\mathcal{C}) - a\) such that 
\(2x \in \pi(\mathcal{C}) - a\) (so that we do not have a sum in \(\mathcal{C}\) when taking two numbers in the same column), 
the colors \(1, ..., n\) are sum-free.

Setting \(w = l = \left\lceil \displaystyle \frac{q}{2} \right\rceil\) and \(r = \left\lfloor \displaystyle \frac{q}{2} \right\rfloor\) satsfies these 
constraints: this is the construction corresponding to Theorem \ref{thm:WS-ab}. \\

By setting \(q = \WS (n)\) and \(p = S(k)\) in Theorem \ref{thm:WS-ab}, one obtains the following corollary.

\begin{corollary}
\label{cor:ineqWS-S}
\(\WS (n+k) \geqslant S(k) \left (\WS (n) + \left \lceil \displaystyle \frac{\WS (n)}{2} \right \rceil +1 \right) + \WS (n)\)
for all \((n,k) \in (\mathbb{N}^*)^2\).
\end{corollary}

This can be seen as an equivalent for weak Schur numbers of Abott and Hanson's construction for Schur numbers 
\cite{AbbottHanson}. This formula also includes the results of Rowley \cite{RowleyWS} as a special case. For \(n>2\), 
the formula from corollary \ref{cor:ineqWS-S} does not give new lower bounds.

\begin{remark}
The inequality from Corollary \ref{cor:ineqWS-S} can be improved by adding 1 to the lower bound if \(\WS (n)\) is odd (more generally if \(q\) is
odd in Theorem \ref{thm:WS-ab}). However, it lengthens the proof and it is never useful in practice.
\end{remark}
\par
Now, we introduce WS-templates and the sequence \(\WS^+\) in order to generalize the above construction.

@@

Note that the special color \(n\) is not necessarily the last color by order of appearance, any color can play this role.
Now, we introduce the number \(\WS^+(n)\) that plays the same role as \(S^+(n)\) for S-templates.

@@

Having established these properties, we prove our main result.

@@

Theorem \ref{thm:Stemp} describes a manner to extend a S-template into a larger Schur partitions using other Schur 
partitions. Likewise, a WS-template can be extended into larger weak Schur partitions by using Schur partitions. This 
is the object of the following theorem.

\begin{theorem}
\label{thm:WStemp}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there are a partition of \([\![1,p]\!]\)
into \(k\) sum-free subsets and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there is a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

\begin{proof}[\textsc{Proof.}]
\begin{sloppypar}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) and \((p,k) \in (\mathbb{N}^*)^2\). Let \(f\) and \(g\) be colorings associated 
to the \(b\)-WS-template and to the sum-free partition of \([\![1,p]\!]\), respectively: 
\({f : [\![1, a + b]\!] \longrightarrow [\![1,n+1]\!]}\) and \({g : [\![1, p]\!]  \longrightarrow [\![1, k]\!]}\). Moreover, 
we assume that the sum-free coloring of \([\![1, p]\!]\) is ordered.
\end{sloppypar}

@@

We split \([\![1, p a + b]\!]\) into three subsets. \\
NB: To keep the notation short, the restriction to \([\![b + 1, p a + b]\!]\) of \(\pi\) defined in Subsection \ref{DefinitionWS+}
is denoted by \(\pi\) in the hereunder equations.

@@

We define a coloring \(h\) of \([\![1, p a + b]\!]\) using \( [\![1,n+k]\!]\) colors:
\[
\begin{array}{c c c l}
	h : & [\![1, p a + b]\!] & \longrightarrow & [\![1,n+k]\!] \\
	& x & \longmapsto &
	\left\{ \begin{array}{l l}
		f(x) & \text{if}~x \in \mathcal{T} \\
		f(\pi(x)) & \text{if}~x \in \mathcal{C} \\
		n + g(\lambda(x)) & \text{if}~x \in \mathcal{R}
	\end{array} \right.
\end{array}
\]

Function \(h\) is well defined since sets \(\mathcal{T}, \mathcal{C}, \mathcal{R}\) partition \([\![1, p a + b]\!]\).
Now, we prove that \(h\) is a weakly sum-free coloring. Let \(x,y \in [\![1, p a + b]\!]\) be such that \(x \neq y\),
\(h(x) = h(y)\), and \(x+y \leqslant p a+ b\). We claim that \(h(x+y) \neq h(x)\). Nine cases are to be distinguished
according to which of the sets \(\mathcal{T}, \mathcal{C}, \mathcal{R}\) \(x\) and \(y\) belong to. It is
sufficient to check only six cases out of nine since \(x\) and \(y\) play symmetric roles. \\

@@

\noindent \underline{\textbf{Case 2:} \((x,y) \in \mathcal{T} \times \mathcal{C}\)}
\par
Given that \(h(x) = h(y)\) and by definition of \(h\), \(f(x) = f(\pi(y))\). Besides, \(f(\pi(y)) \leqslant n\) since
\(y \in \mathcal{C}\). Two cases are to be distinguished according to the value of \(x + \pi(y)\).
\begin{itemize}
\item \begin{sloppypar}
	If \(x + \pi(y) \leqslant a + b\) then \(f(x + \pi(y)) = f(\pi(x + y))\) (Proposition \ref{prop3}). Given that \(f\) is
	a weakly sum-free coloring, \(f(x+\pi(y)) \neq f(x)\) since \(f(x)=f(\pi(y))\) and \(x \neq \pi(y)\) since
	\({x \leqslant b < \pi(y)}\).
	\end{sloppypar}
\item \begin{sloppypar}
	If \(x+\pi(y)> a+b\) then given that \(f\) is a WS-template and since \({f(x) = f(\pi(y)) \leqslant n}\),
	\({f(\pi(x+\pi(y))) \neq f(x)}\).~Furthermore \({f(\pi(x+\pi(y))) = f(\pi(x+y))}\) (Proposition \ref{prop2}), such that
	\({f(\pi(x+ y)) \neq f(x)}\).
	\end{sloppypar}
\end{itemize}
\par
In both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

\noindent \underline{\textbf{Case 3:} \((x,y) \in \mathcal{T} \times \mathcal{R}\)}
\par
Necessarily \(h(x) = h(y) = n + 1\). Two cases are to be distinguished according to the value of \(\lambda(x+y)\).
\begin{itemize}
\item If \(\lambda(y)=\lambda(x+y)\) then \(\pi(x + y) = x + \pi(y)\) (Proposition \ref{prop6}). By definition of
	\(h\), \(f(x) = f(\pi(y))\). Given that \(f\) is a weakly sum-free coloring, \(f(x + \pi(y)) \neq f(x)\) since
	\(f(x) = f(\pi(y))\) and \(x \neq \pi(y)\) since \({x \leqslant b < \pi(y)}\). Hence \(h(x + y) \neq h(x)\).
\item If \(\lambda(y) \neq \lambda(x + y)\) then \(\lambda(x + y) = \lambda(y) + 1\) since \(x \leqslant b < a\).
	Besides, \(n + 1 = h(y) = n +  g(\lambda(y))\). Hence \(g(\lambda(y)) = 1\). Moreover \(g(1) = 1\) since \(g\)
	is an ordered coloring. Therefore, given that \(g\) is sum-free, \(g(\lambda(y) + 1) \neq 1\). If \(\pi(x + y) \in
	A_{n + 1}\) then \(h(x + y) = n + g(\lambda(x + y)) \neq n + 1\). Otherwise, \(h(x + y) \leqslant n\).
\end{itemize}
\par
In both cases \(h(x + y) \neq h(x)\).

\noindent \underline{\textbf{Case 4:} \((x,y) \in \mathcal{C}^2\)}
\par
By definition of \(h\) and since \(h(x)=h(y)\), \(f(\pi(x)) = f(\pi(y))\). Two cases are to be distinguished according
to the value of \(\pi(x)+\pi(y)\).
\begin{itemize}
\item If \(\pi(x) + \pi(y) \leqslant a+b\) then \(\pi(x)+\pi(y) = \pi(x + y)\). Hence \(f(\pi(x + y)) \neq f(\pi(x))\) since
	\(f\) is sum-free for \(x>b\)
\item \begin{sloppypar}
	If \(\pi(x)+\pi(y)>a+b\) then given that \(f\) is a WS-template, \({f(\pi(\pi(x)+\pi(y))) \neq f(\pi(x))}\) since
	\({f(\pi(x)) = f(\pi(y))}\). Besides,  \({f(\pi(\pi(x)+\pi(y))) = f(\pi(x + y))}\) (Proposition \ref{prop4}). Hence \({f(\pi(x + y))
	\neq  f(\pi(x))}\).
	\end{sloppypar}
\end{itemize}
\par
In both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

@@

Theorem \ref{thm:WS-ab} can be seen as a particular case of Theorem \ref{thm:WStemp} in the same way Abott and Hanson's 
construction \cite{AbbottHanson} can be seen as a particular case of theorem \ref{thm:Stemp}.

@@

By setting \(p = S(k)\) and \(a = \WS^+(n + 1)\) in Theorem \ref{thm:WStemp}, one obtains:

\begin{remark}
In the S-template construction for Schur numbers, the additive constant comes from the fact that the special color does
not necessarily appear at the very begining of the repeating pattern. Likewise, \(b_{max}\) can actually be replaced by
\[
\max_{b \in \mathbb{N}^*} \left\{\min (A_{n+1} \backslash [\![1, b ]\!]) - 1~|~ \WS_b^+(n+1) = \WS^+(n+1) \right\}.
\]
\end{remark}

@@

\begin{proposition}
\label{prop:tunelastrow}
Let \((b, k, a) \in (\mathbb{N}^*)^3\) and let \(f\) be a coloring associated to a \(b\)-WS-template with width \(p\) and \(k\) colors. Let
\(c \in \mathbb{N}\) and assume there there exists a coloring \(g\) of \([\![b + 1, b + c]\!]\) with \(k\) colors such that for all \(c \in [\![1, k]\!]\),

@@

The WS-templates can actually be fine-tuned further. However, it only gives minor improvements (most likely only an 
additive constant) while increasing the size of the search space.

\par
These modifications rely on one main observation: the first row (excluding the "tail") has constraints that other rows do not 
have because of the tail, especially if the special color appears in the tail as well. Therefore, having a coloring dedicated to 
the first row and another coloring dedicated to the repeating pattern would weaken the constraints on the repeating pattern.

\par
These additionnal constraints on the first row imply additionnal constraints on the rows of the its color (compared to other 
colors). However, one may notice that there are never two consecutive rows of this color. Moreover these additionnal constraints 
mostly concern two consecutive rows. Having a special coloring for rows of this color as well as another color dedicated to the 
first row avoids enforcing these constraints specific to one color on the other colors and thus on the main repeating pattern.

\par
Finally, given that the color of the first row is different from the special color of the main repeating pattern, coloring \(g\) in
Proposition \ref{prop:tunelastrow} can exploit this difference. \\

\begin{proof}[\textsc{Proof.}]
The idea is the same as in theorem \ref{thm:WStemp}. The properties specific to WS-templates are inherited 
from both the S-template and the WS-template. \\
\end{proof}

@@

We produced WS-templates using the lingeling SAT solver \cite{Lingeling2017}, hence providing lower bound on \(WS^+\) 
and inequalities of the type \(WS(n+k) \geqslant a S(n) + b\). We sought templates providing the greatest value of \((a, b)\) 
(in the lexicographic order). Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
Here are the inequalities given by the current best WS-templates. The \hyperref[WS-templates]{template} 
corresponding to the third inequality can be found in the \ref{WS-templates}.

@@

However, trying out all of the 2\,447\,113\,088 Schur number 5-partitions \cite{Heule2017} one by one would not result in a 
reasonable computation time: it is necessary to test the construction on several partitions at once. We encoded the problem as a 
satisfiability problem and used the 1616 backdoors that were used in \cite{Heule2017} in order to encode a group of partitions in 
a compact and efficient way. Among all of the 1616 backdoors, only the 911\textsuperscript{th} backdoor gave a weakly sum-free 
partition of length 643. This backdoor gave a weakly sum-free partition of length 646 and cannot give a weakly sum-free 
partition of length 647. A weakly sum-free partition of length 646 can be found in the \ref{WS(6)}.
